PLUNGER: Reproducing FLUSH+RELOAD: A High-Resolution, Low-Latency Side Channel Attack On GnuPG

### 2014-12-17 paper/paper.tex
Fixed some misc formatting and wording.
Also extended the section on page sharing.

A typical cache architecture is laid out in three levels, generally referred to
as the \textit{L1}, \textit{L2}, and \textit{L3}. The L3 cache is usually the
largest of the three, then the L2, and finally the L1; L1 and L2 caches are also
generally allocated on a per-core basis in the case of multi-core CPUs, whereas
the L3 cache is usually shared between all cores, as it is the largest. L1
caches are also typically divided into \textit{instruction} sections and
\textit{data} sections, each reserved for that particular type of instruction.
It is equally important for instruction memory as well as data memory to be
cached, as both require fast access in order for the CPU to be performant. For
example, the Intel Haswell i7 4600U CPU has a 128KiB L1 cache, divided into two
64KiB caches for instructions and data; a 512KiB L2 cache, and a 4MiB L3 cache.

@@

Windows, Linux, and possibly OS X \footnote{It is unclear whether or not OS X
implements page deduplication, but as far as we can tell it does} also perform a
more aggressive form of page sharing known as \textit{page deduplication}. With
this scheme, the operating system scans memory pages during idle CPU cycles, and
looks for pages that are identical in content. It then collapses them into one
singular page that is then shared by all processes that are accessing the
content of that particular page. As a result of this page deduplication scheme,
processes may be unknowingly or unintentionally sharing memory with each other.
However, to prevent malicious access to the shared pages, the operating system
maps shared pages as copy-on-write; that is, when a process wishes to write to a
page, it must first make a local copy for itself that it can safely modify. When
a write is requested, the CPU traps to the kernel, which then remaps a copy of
the desired page for the process requesting the write, and then returns control
to the calling process. Read operations in this scheme are safe, as they do not
modify the targeted page.

@@

The \flushreload{} attack takes advantage of page deduplication on modern
processors. The basic attack works as follows. The attacker loads into memory
some block of data that it believes the victim to also be accessing. The system
then deduplicates these two identical pages. The attacker can then use the
instruction \texttt{clflush} to flush this from its $L1$ cache, which flushes it
from the $L3$ cache, and consequently the $L1$ cache of the victim as well. The
attacker then waits for some time slot $T$, and attempts to access a cache line
in this page, and measures the time (in cycles) it takes for the line to be
loaded. If the cache line was accessed by the victim during $T$, the time taken
will be shorter, and if it was not accessed, the time taken will be longer.
Consequently, the attacker can determine if the line was accessed during $T$.

@@

We also attempted to enforce the fixed slot by dynamically adjusting the busy
wait time based on how long all of the probes took. While, in theory, this
technique would yield better results by ensuring that we take cache timing
readings uniformly, in practice, this did not yield significantly better
results.

@@

Because the particular \flushreload{} side channel vector, the modular
exponentiation function, that we (and Yarom and Falkner\cite{YF13}) were
targeting was mitigated in GPG 1.4.14, we used GPG 1.4.12. We built GPG with the
default compilation flags, which included building with \texttt{-O2}
optimizations and debug flags.

@@

Arch Linux provided the poorest results by far. No meaningful data was
discernible under any configuration, as there is timing noise for the entire
duration of spying, which in the case of the figure below, occurred roughly
between timeslots 7000 and 18000 on the x-axis.. Furthermore, it was fairly
opaque as to why this was occurring, as there were no significant differences
between this and the configuration running on OS X. The largest obvious
difference between the two systems is that GnuPG on Arch Linux was compiled
under GCC 4.9.2, as opposed to GCC 4.2 on OS X; however, it is not clear how
this would change the output this drastically, or if it did at all. Two
additional hypotheses for why this behaviour was exhibited were that Arch Linux
was being run as a guest OS on a Windows 8.1 host, and that it was being run on
a Haswell-series Intel i7, which might require compiling the GnuPG binary
differently and therefore might exhibit different behaviour, as this CPU series
did not exist at the time of the original paper's publishing.  However, both of
these hypotheses were disproven by testing on a second machine running Arch
Linux as a host OS on an older Intel Celeron CPU, in which the attack produced
the same noisy results.

@@

Many of our preliminary results were plagued by a lot of noise, a lot of which
resulted from speculative execution. To reduce the effects of this, we attempted
to probe the very close to the ends of the respective functions. We found that
simply setting breakpoints to the relevant lines in GDB would not yield good
addresses to probe. Even setting a breakpoint to the last line of the function
in the source would not return an address of an instruction close to the return.
To find the best addresses, we disassembled the GPG binary and located addresses
near each function's \texttt{retq} instruction.

>>>>

A typical cache architecture is laid out in three levels, generally referred to
as the \textit{L1}, \textit{L2}, and \textit{L3}. The L3 cache is usually the
largest of the three, then the L2, and finally the L1; L1 and L2 caches are also
generally allocated on a per-core basis in the case of multi-core CPUs, whereas
the L3 cache is usually shared between all cores, as it is the largest. L1
caches are also typically divided into \textit{instruction} sections and
\textit{data} sections, each reserved for that particular type of instruction.
It is equally important for instruction memory as well as data memory to be
cached, as both require fast access in order for the CPU to be performant. For
example, the Intel Haswell i7 4600U CPU has a 128KiB L1 cache, divided into two
64KiB caches for instructions and data; a 512KiB L2 cache; and a 4MiB L3 cache.

@@

Windows, Linux, and possibly OS X \footnote{It is unclear whether or not OS X
implements page deduplication, but as far as we can tell it does} also perform a
more aggressive form of page sharing known as \textit{page deduplication}. With
this scheme, the operating system scans memory pages during idle CPU cycles, and
looks for pages that are identical in content. It then collapses them into one
singular page that is then shared by all processes that are accessing the
content of that particular page. As a result of this page deduplication scheme,
processes may be unknowingly or unintentionally sharing memory with each other.
However, to prevent malicious access to the shared pages, the operating system
maps shared pages as copy-on-write; that is, when a process wishes to write to a
page, it must first make a local copy for itself that it can safely modify. When
a write is requested, the CPU traps to the kernel, which then remaps a copy of
the desired page for the process requesting the write, and then returns control
to the calling process. Read operations in this scheme do not require such a trap
as they do not modify the targeted page and are therefore ``safe''.

@@

The \flushreload{} attack takes advantage of page deduplication on modern
processors. The basic attack works as follows: the attacker loads into memory
some block of data that it believes the victim to also be accessing. The system
then deduplicates these two identical pages. The attacker can then use the
instruction \texttt{clflush} to flush this from its $L1$ cache, which flushes it
from the $L3$ cache, and consequently the $L1$ cache of the victim as well. The
attacker then waits for some time slot $T$, and attempts to access a cache line
in this page, and measures the time (in cycles) it takes for the line to be
loaded. If the cache line was accessed by the victim during $T$, the time taken
will be shorter, and if it was not accessed, the time taken will be longer.
Consequently, the attacker can determine if the line was accessed during $T$.

@@

We also attempted to enforce the fixed slot by dynamically adjusting the busy
wait time based on how long all of the probes took. While in theory this
technique should yield better results by ensuring that we take cache timing
readings uniformly, in practice it did not yield significantly better
results.

@@

Because the particular \flushreload{} side channel vector, the modular
exponentiation function, that we (and Yarom and Falkner\cite{YF13}) were
targeting was mitigated in GPG 1.4.14, we used GPG 1.4.12. We built GPG 
using GCC with the default compilation flags, which included
building with \texttt{-O2} optimizations and debug flags.

@@

Arch Linux provided the poorest results by far. No meaningful data was
discernible under any configuration, as there is timing noise for the entire
duration of spying, which in the case of the figure~\ref{fig:data-arch}, occurred roughly
between timeslots 7000 and 18000 on the x-axis.. Furthermore, it was fairly
opaque as to why this was occurring, as there were no significant differences
between this and the configuration running on OS X. The largest obvious
difference between the two systems is that GnuPG on Arch Linux was compiled
under GCC 4.9.2, as opposed to GCC 4.2 on OS X; however, it is not clear how
this would change the output this drastically, or if it did at all.

Two additional hypotheses for why this behaviour was exhibited were that Arch Linux
was being run as a guest OS on a Windows 8.1 host, and that it was being run on
a Haswell-series Intel i7, which might require compiling the GnuPG binary
differently and therefore might exhibit different behaviour, as this CPU series
did not exist at the time of the original paper's publishing.  However, both of
these hypotheses were disproven by testing on a second machine running Arch
Linux as a host OS on an older Intel Celeron CPU, in which the attack produced
the same noisy results.

Additionally, it may be the case that the addresses determined with GDB were inaccurate,
and that probing based on addresses determined via \texttt{objdump} would produce
better results.

@@

Many of our preliminary results were plagued by a lot of noise, a lot of which
resulted from speculative execution. To reduce the effects of this, we attempted
to probe the very close to the ends of the respective functions. We found that
simply setting breakpoints to the relevant lines in GDB would not yield good
addresses to probe. Even setting a breakpoint on the last line of the function
in the source would not return an address of an instruction close to the return.
To find the best addresses, we disassembled the GPG binary and located addresses
near each function's \texttt{retq} instruction.
